Dispersion Distance and the Matter Distribution of the Universe in Dispersion Space

### 2015-06-04 writeup/dm_space_distortions_vKris1_Jun3.tex
A bunch of edits to intro and abstract and title

\title{The large-scale structure of the Universe in dispersion measure space}

@@

    The recently discovered fast radio bursts (FBRs) are a new class of radio
    transient that, due to their high dispersion from intervening plasma,
    have been interpreted as
    originating from cosmological distances.  The rate of fast radio bursts
    has been estimated to be several thousand over the whole sky per day.  If
    cosmological, the sources of these events should trace the large-scale
    structure of the
    Universe.
    We show here that, even without redshifts, these bursts can be used to
    study the three-dimensional cluster of matter in the Universe.
    This is because the dispersion of radio waves as they travel through the
    Universe, like redshift, can be
    used as a proxy for radial distance. 
    Due to inhomogeneities in
    the plasma's density, dispersion is an imperfect proxy for radial
    distance and we show that this leads to calculable dispersion-space
    distortions in the apparent clustering. We calculate the dispersion-space
    power spectra for a simple model of electron and FRB clustering and show that
    the clustering signal could be detected in a survey of $10^4$ FRB
    events. This rate is \ldots as is expected to be produced by the Canadian Hydrogen Intensity
    Mapping Experiment.

\end{abstract}

@@

However, redshift is not the only measure of radial distance. In particular,
dispersion of electromagnetic signals due to intervening plasma
is readily measured from fast radio bursts (FRB) which are a class
of radio transient first reported by \citet{2007Sci...318..777L}.
There have since been 
10 such bursts reported
\citep{2012MNRAS.425L..71K,2013Sci...341...53T,
2014ApJ...790..101S, 2014ApJ...792...19B, 
2015MNRAS.447..246P,2015ApJ...799L...5R} and total rate of bursts
been estimated to be $3.3^{+5}_{-2.5} \times 10^3$
per sky per day \cite{2015arXiv150500834R}. 
The dispersion measure
(DM) is a line-of-site integral of the free electron density and can thus be
used as a distance estimate, as is routinely done with galactic pulsars.
FRB events are observed to have dispersion measures of order
$1000\,{\rm pc} / {\rm cm}^3$, greatly in excess of
the galactic expectation of order $100\,{\rm pc} / {\rm cm}^3$, depending on
galactic latitude
\cite{2014arXiv1412.4829D}.
They must therefore be of extra-galactic origin, with their
dispersion produced either by the
inter-galactic medium (IGM) \citep{2013Sci...341...53T} or the
environment of the source
\citep{2014ApJ...785L..26L,2015arXiv150101341P,2015arXiv150505535C}.
If the dispersion is dominated by the inter-galactic medium,
then the average DM--distance relation can be modelled
\citep{2014ApJ...797...71Z,2004MNRAS.348..999I, 2003ApJ...598L..79I}
and the sources are at
cosmological distances of a few gigaparsecs. This would permit
the study of the clustering of FRB sources in three dimensions which should trace the
large-scale structure of the Universe on linear scales. Note that although we
frame our discussions in terms of FRBs, dispersion may be

>>>>

%\title{The large-scale structure of the Universe in dispersion measure space}
\title{Dispersion Distance and the Matter Distribution of the Universe in Dispersion Space}

@@

    We propose that `standard pings', brief broadband radio impulses, can be used to
    study the three-dimensional clustering of matter in the Universe even in the absence of redshift information.
    The dispersion of radio waves as they travel through the
    Universe, like redshift, can be used as a cosmological distance measure.
    Due to inhomogeneities in the electron density along the line-of-siight dispersion is an imperfect proxy for radial
    distance and we show that this leads to calculable dispersion-space
    distortions in the apparent clustering of sources. Fast radio bursts (FRBs) are a new class of radio
    transients that are the prototypical standard ping, and due to their high observed dispersion from intervening plasma,
    have been interpreted as originating at cosmological distances.  The rate of fast radio bursts
    has been estimated to be several thousand over the whole sky per day, and if cosmological, the sources of these events should trace the large-scale
    structure of the Universe.  We calculate the dispersion-space
    power spectra for a simple model where electrons and FRBs are biased tracers of the large-scale-structure of the Universe and show that
    the clustering signal may be detectable in a survey with as few as $10000$ 
    events.  Upcoming wide-field radio telescopes, like the Canadian Hydrogen Intensity Mapping Experiment (CHIME), are expected to detect...
\end{abstract}

@@

However, redshift is not the only measure of radial distance.   
Standard candles and standard rulers can be used to estimate distance using brightness and angular size respectively.
In the same way a \emph{standard ping}, a brief broadband radio impulse, can be used to estimate radial distance because of
the characteristic $\lambda^2$ dispersion of electromagnetic signals along the line-of-sight as a function of wavelength.
This technique, widely used as an approximate distance indicator to pulsars in the Galaxy, can be used to define the
\emph{dispersion distance} to cosmological sources.

due to intervening plasma
is readily measured from fast radio bursts (FRB) which are a class
of radio transient first reported by \citet{2007Sci...318..777L}.
There have since been 
10 such bursts reported
\citep{2012MNRAS.425L..71K,2013Sci...341...53T,
2014ApJ...790..101S, 2014ApJ...792...19B, 
2015MNRAS.447..246P,2015ApJ...799L...5R} and total rate of bursts
been estimated to be $3.3^{+5}_{-2.5} \times 10^3$
per sky per day \cite{2015arXiv150500834R}. 
The dispersion measure
(DM) is a line-of-site integral of the free electron density and can thus be
used as a distance estimate, as is routinely done with galactic pulsars.
FRB events are observed to have dispersion measures of order
$1000\,{\rm pc} / {\rm cm}^3$, greatly in excess of
the galactic expectation of order $100\,{\rm pc} / {\rm cm}^3$, depending on
galactic latitude
\cite{2014arXiv1412.4829D}.
They must therefore be of extra-galactic origin, with their
dispersion produced either by the
inter-galactic medium (IGM) \citep{2013Sci...341...53T} or the
environment of the source
\citep{2014ApJ...785L..26L,2015arXiv150101341P,2015arXiv150505535C}.
If the dispersion is dominated by the inter-galactic medium,
then the average DM--distance relation can be modelled
\citep{2014ApJ...797...71Z,2004MNRAS.348..999I, 2003ApJ...598L..79I}
and the sources are at
cosmological distances of a few gigaparsecs. This would permit
the study of the clustering of FRB sources in three dimensions which should trace the
large-scale structure of the Universe on linear scales. Note that although we
frame our discussions in terms of FRBs, dispersion may be
